---
title: "Wrapping Spotify Wrapped"
author: "Ann Basting"
date: "2023-01-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(spotifyr)
library(knitr)
library(corrr)
library(shadowtext)
library(jsonlite)
library(lubridate)
library(gghighlight)
library(tidyverse)
library(plotly)
library(heatmaply)

source("../workbook/20221222/alex_fxns.R")
source('../workbook/authentication.R')
```

## Spotify's API

While I was waiting for Spotify to send me my data, I started playing around with Spotify's API by using the `spotifyr` package. 

### Vizualizing Genre

The `spotifyr` package allows you to pull up to 50 of your top artists and the genres associated with each of them. For this first part, I was curious about which genres appeared most frequently among my 50 most listened to artists.

```{r API, echo = FALSE, fig.dim = c(8, 9.5)}
#get top artists
top_artists <- get_my_top_artists_or_tracks(type = 'artists', 
                                            time_range = 'long_term', 
                                            limit = 50, 
                                            authorization = get_spotify_authorization_code()
) %>% 
  select(name, genres) %>% 
  rowwise %>% 
  mutate(genres = paste(genres, collapse = ', ')) %>% 
  ungroup %>%
  kable()

#only extract top genres from my top artists

top_genres <- get_my_top_artists_or_tracks(type = 'artists', 
                                           time_range = 'long_term', 
                                           limit = 50, 
                                           authorization = get_spotify_authorization_code()
) %>% 
  select(genres) %>%
  unlist() 

top_genres <- as.data.frame(table(top_genres)) #get counts

#only include genres that appear more than once in dataset
# top_genres <- top_genres %>%
#   filter(Freq > 1) 

#order by decreasing frequency
top_genres <- top_genres[order(top_genres$Freq, decreasing = T), ]

#generate plot
png(filename="genre_freq.png", width=600, height=1000)
genre_plt <- top_genres %>%
  mutate(top_genres = fct_reorder(top_genres, Freq)) %>% #reordering top genres by frequency
  ggplot() +
  geom_col(aes(Freq, top_genres), fill = "#1DB954", width = 0.6) + 
  scale_x_continuous(
    limits = c(0, 17),
    breaks = seq(0, 17, by = 5), 
    expand = c(0, 0), # The horizontal axis does not extend to either side
    position = "top"  # Labels are located on the top
  ) +
  # The vertical axis only extends upwards 
  scale_y_discrete(expand = expansion(add = c(0, 0.5))) +
  theme(
    # Set background color to white
    panel.background = element_rect(fill = "white"),
    # Set the color and the width of the grid lines for the horizontal axis
    panel.grid.major.x = element_line(color = "#A8BAC4", size = 0.3),
    # Remove tick marks by setting their length to 0
    axis.ticks.length = unit(0, "mm"),
    # Remove the title for both axes
    axis.title = element_blank(),
    # Only left line of the vertical axis is painted in black
    axis.line.y.left = element_line(color = "black"),
    # But customize labels for the horizontal axis
    axis.text.x = element_text(#family = "Econ Sans Cnd", 
      size = 10)
  ) +
  ggtitle('Frequency of Genres Associated with my Top 50 Spotify Artists')

ggplotly(genre_plt)

```

### Song Features

I also discovered that Spotify's API provides interesting metrics on all songs (ex. valence, danceability, liveness). 

#### Average

First things first, I wanted to see the average of these metrics for all 2,000+ songs in one of my playlists (aka Tune Void).

```{r feature, echo=FALSE}
#read in my playlists
user_id <- 'victorialokelani'
playlists <- get_user_playlists(user_id,
                                limit = 20,
                                offset = 0,
                                authorization = get_spotify_authorization_code())
#pull id from tune void playlist
playlist_uris <- playlists %>%
  filter(name == "tune void (est. 2014)") %>%
  select(id) %>%
  pull(id)

#pull in audio features of tune void
tune_void <- get_playlist_audio_features(user_id, playlist_uris)

#keep these columns
tune_void_columns_to_keep <- c("danceability", "energy", "key", "loudness", 
                               "mode", "speechiness", "acousticness", 
                               "instrumentalness", "liveness", "valence", "tempo", 
                               "time_signature", "track.artists", 
                               "track.duration_ms", "track.name", "track.popularity", 
                               "track.album.name", "track.album.release_date", 
                               "key_name", "mode_name", "key_mode")

#create playlist with these columns
tune_void2 <- tune_void %>%
  select(all_of(tune_void_columns_to_keep))

#find average of num values
tune_void2 %>% 
  select(c(1:2, 4, 6:11)) %>%
  colMeans()
```

#### Correlation Network

Next, I wanted to see how the features of the songs in this playlist related to each other. Correlation networks look cool, so I went with that.

```{r corrnet, echo=FALSE}
#create a correlation network
cor <- correlate(tune_void2[,c(1:2, 4, 6:11)])
network_plot(cor, min_cor = 0.01)
```

#### Heatmap

Finally, I wanted to take the 50 most recently added songs from the Tune Void and see how those songs cluster together based on the metrics Spotify provides.

```{r heatmap, echo = FALSE, fig.dim = c(9, 9.5)}
# turn song names into row names (only 50 most recent songs added)
tune_void_norm <- as.data.frame(tune_void2)
songs <- tail(tune_void_norm[,15], n = 50)
tune_void_norm <- tail(scale(tune_void_norm[,c(1:2,4,6:11)]), n = 50)
rownames(tune_void_norm) <- songs

# subset last 100 rows of tune void and normalize data
tune_void_norm <- as.matrix(tune_void_norm)


# trying to build a heatmap of 50 most recent songs added to tune void
hmap <- heatmaply(
          tune_void_norm,
          k_col = 2, 
          k_row = 2
        )
hmap
```
## One Year of Streaming Data

[write later]

### Alex's Code

#### Tracks by Release Date

```{r tracks by release date, echo=FALSE}
top50lt <- get_my_top_artists_or_tracks(
  type = 'tracks', time_range = 'long_term', limit = 50
) %>%
  mutate(artist.name = map_chr(artists, function(x) x$name[1])) %>%
  mutate(duration_ms = as.numeric(duration_ms)) %>%
  mutate(duration_s = duration_ms/1000, duration_m = duration_ms/60000)

top50lt$rank <- 1:nrow(top50lt)

## Plotting the years since track release by rank
## Creating a variable that codes for how many days it's been since the track
## was released
top50lt$album.release_year <- substr(top50lt$album.release_date, 1, 4)
top50lt$album.release_year <- as.numeric(top50lt$album.release_year)
top50lt$years_since_release <- as.numeric(year(today()))-top50lt$album.release_year

## I have a song that was released this year, so I'm going to replace 0s in the 
## years_since_release with 0.5 and see if that shows up in plotly
top50lt$years_since_release <- ifelse(top50lt$years_since_release == 0, 
                                      0.5, top50lt$years_since_release)

ggplotly(
  ggplot(top50lt, aes(x = rank, y = years_since_release, fill = artist.name)) +
    geom_bar(stat = 'identity') +
    theme_light() +
    theme(legend.position = 'none') +
    ylab("years since track release") +
    ggtitle("Years Since Track Release by Rank")
)

```

```{r more alex functions, echo=FALSE}
# import streaming data and clean
myStreamingData <- import_spotify_streaming_data(datafile = "../data")

myCleanData <- clean_spotify_streaming_data(data = myStreamingData,
                                            your_timezone = "EST")

# plot!!!!
plot_streaming_timeofday(myCleanData, artist_cutoff = 10)

plot_streaming_artists(myCleanData, artist_cutoff = 10)
```